---
title: "R_HandBook"
author: "Vishal Kumar"
date: "September 28, 2017"
output: html_document
---
This handbook contains very bacsis of R programming to advanced level syntax with the example application. Its only created for personel learning.

These are the normal commands in R to operate on different things. These are just for showing sample of a R session.

```{r sample_of_R_session}
age <- c(6,7,9,4,7,3,2,5,7,8,4,4,7,3,7)
weight <- c(22, 32.9, 23, 54, 23.7, 23, 43, 23.3, 56, 76, 23.5, 12.5, 23, 34, 45)
mean <- mean(weight)
standard_deviation <- sd(weight)
correlation <- cor(age, weight)
plot(age, weight)
q()
```

Now you can again open your R command window.

Codes for working with the packages in R.

```{r packages_stuff}
install.packages("vcd")
library(vcd)
help(Arthritis)
Arthritis
```


*Data Structures in R
+Vectors
Vectors are one dimensional array that are used to hold the variables i.e numeric variable , logical variable, character variable.
```{r Vectors}
a <- c(2, 5, 6, 49)
b <- c("vishal", "vivek",  "neha")
c <- c(TRUE, FALSE, FALSE, TRUE)
a[3]
b[2:3]
a[c(2,3,4)]
```

+Matrices
A matrix is a two dimensional array where each element have same node(numeric, character etc.)

```{r Matrix}
mat <- matrix(c(1:20), nrow = 5, ncol = 4)
mat[2,]
mat[, 4]
mat[1, c(2,4)]
cells <- c(1, 26, 24, 68)
rnames <- c("R1", "R2")
cnames <- c("C1", "C2")
mymatrix <- matrix(cells, nrow = 2, ncol = 2, byrow = TRUE, dimnames = list(rnames, cnames))
mymatrix

```


+Arrays
Arrays are similar to matrices but can have more than two dimensional. They are created with an array function. 
 
```{r Arrays}
dim1 <- c("A1", "Á2")
dim2 <- c("B1", "B2", "B3")
dim3 <- c("C1", "C2", "C3", "C4")
arr <- array(1:20, c(2,3,4), dimnames = list(dim1, dim2, dim3) )
arr
```


+Data Frames
Data frames are the much simpler structure than matrices. It contain rows and coloums and different coloumn of data frame can contain different types of data frames.

```{r Data_Frames}
patientId <- c(1,2,3,4)
age <- c(23,32,12,14)
diabetes<- c("type1", "type2", "type2", "type1")
status <- c("improved", "same", "poor", "improved")
patientData <- data.frame(patientId, age, diabetes, status)
patientData
patientData[2,4]
patientData[, 3]
```



+Factors
Factors are 





#Simple linear regression
In simple linear regression we fit a linear models and predict response variable from predicting variable.

```{r linear_regression}
fit <- lm(weight ~ height, data = women)
summary(fit)
women$weight
fitted(fit)
plot(women$height, women$weight, xlab = "weigth of womens", ylab = "height of womens")
abline(fit)

```


#Polynomial regression
We can fit quadratic equation to the model to improve our model.

```{r ploynomial_regression}
fit2 <- lm(weight ~ height + I(height^2), data = women)
summary(fit2)
plot(women$height, women$weight, xlab = "WH", ylab = "WW")
lines(women$height, fitted(fit2))
```


We can also fit cubic equation to our models.

```{r cubical_regression}
fit3 <- lm(weight ~ height + I(height^2) + I(height^3), data = women)
summary(fit3)
plot(women$height, women$weight, xlab = "WH(in inches)", ylab = "WW(in kgs)")
lines(women$height, fitted(fit3))
```


#Multiple Regression
In multiple regression we more than one predictor variables for the prediction of respomse variable.

```{r Multiple_regression}
fit <- lm(mpg ~ hp + wt, data = mtcars)
summary(fit)

```

```{r mul_lin_reg_withInteraction}
fit <- lm(mpg ~ hp + wt + hp:wt, data = mtcars)
summary(fit)
library(effects)
plot(effects("hp:wt", fit, list(wt = c(2.2, 3.2, 4.2))), multiline = TRUE)
```


```{r bagged_loess}
install.packages("ElemSTatLearn")
library(ElemStatLearn)
data(ozone, package = "ElemStatLearn")
ozone <- ozone[order(ozone$ozone), ]
head(ozone)
ll <- matrix(NA, nrow, ncol = 155)
for(i in 1:10){
  ss <- sample(1:dim(ozone)[1], replace = T)
ozone0 <- ozone[ss,]; ozone0 <- ozone0[order(ozone0)$ozone, ]
loess0 <- loess(temprature ~ ozone, data = ozone0, span = 0.2)
ll[i,] <- predict(loess0, newdata = data.frame(ozone = 1:155))
}

```



```{r Bagged_loess_plot}
plot(ozone$ozone, ozone$tempreature, pch = 19, cex = 0.5)
for(i in 1:10){
  lines(1:155, ll[i,], col = "grey", lwd = 2)
}
lines(1:155, apply(ll, 2, mean), col = "red", lwd = 2)

```

```{r bagging_tree}
data("iris")
library(ipred)
bagTree <- bagging(Species ~ ., data = iris, coob = TRUE)
print(bagTree)
bagTree$mtrees[[2]]$btree

```


```{r random_forest_bagging}

library(randomForest)
forestIris <- randomForest(Species ~ Petal.Width+petal.length, data = iris, prox = TRUE)
forestIris
getTree(forestIris, k = 2)


iris.p <- classCenter(iris[, c(3, 4)], iris$Species, forestIris$prox)
plot(iris[,3], iris[, 4], pch = 21, xlab = names(iris)[3], ylab = names(iris)[4],
     bg = c("red", "blue", "green")[as.numeric(factor(iris$Species))],
     main = "Iris  Data Prototypes"
)
points(iris.p[,1], iris.p[,2], pch = 21, cex = 2, bg = c("red", "blue", "green"))
```



